Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-22 shanghai.tex
Made some changes to the notation for Weil group, working on the CS version of Nm

\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\Weil}{W}
\newcommand{\Frob}{{\operatorname{Fr}}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\Irrep}{Irrep}
\DeclareMathOperator{\Pic}{Pic}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Hh}{H}
\DeclareMathOperator{\Res}{Res}

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}

@@

\begin{remark} \label{rem:CS-action}
If $W$ is a group equipped with a homomorphism $\alpha : W \to \Gal(\bFq/\Fq)$
then an \emph{action} of $W$ on $\gcs{L}$ is a collection of isomorphisms
\[
\gamma_w : \alpha(w)^* \gcs{L} \to \gcs{L}
\]
so that $\gamma_{uw} = \gamma_w \circ \alpha(w)^*\gamma_u$. \David{make sure I have the order right and add a reference}
Specifying an $F$ as in \ref{CS.2} is the same as giving an action of the Weil group $W_k$
on $\gcs{L}$ compatible with $\mu$.
\end{remark}

@@

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\Weil(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil(\bFq/\Fq)\to \Weil(G,\bg)$ of the exact sequence
  \[
  \begin{tikzcd}
  1 \rar & \pi_1(\bG, \bg) \rar &  \Weil(G,\bg) \rar
  & \ar[swap,bend right, dotted]{l}{g} \Weil(\bFq/\Fq) \rar & 1
  \end{tikzcd}
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil(\bFq/\Fq)$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = t^\Frob_{(\gcs{L},F)}(g)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil(\bFq/\Fq)$ is generated by
  $\Frob$ and $t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx$ is
  a character, it follows that the image of $t_{(\gcs{L},F)}(g)$ is a
  finite subset of $\mu(\EE)$. Thus, $t_{(\gcs{L},F)}(g)$ extends to
  an $\ell$-adic character of $\Gal(\bFq/\Fq)$, which will be denoted
  by the same symbol.
  \[
  \begin{tikzcd}
  1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \Weil(G,\bg) \rar \dar & \Weil(\bFq/\Fq) \rar \dar & 1\\
  1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1
  \end{tikzcd}
  \]

@@

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil(\bFq/\Fq), \Hom_\text{grp}(G(\bFq),\EEx)),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [t_\cs{L}],
  \]
  where $t_\cs{L}: \Weil(\bFq/\Fq)\to \Hom_\text{grp}(G(\bFq),\EEx)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad t_\cs{L}(\Frob)(g) = t^\Frob_\cs{L}(g).
  \]
\end{proposition}

\begin{proof}
  To simplify notation, we set $W_\Fq \ceq \Weil(\bFq/\Fq)$ and
  $G(\bFq)^* \ceq \Hom_\text{grp}(G(\bFq),\EEx)$.  Recall the
  equivalence between etale group schemes over $\Fq$ and the category
  of groups equipped with a continuous action of $\Gal(\bFq/\Fq)$ (for
  the discrete topology on the group), given by $G \mapsto G(\bFq)$.
  Under this equivalence any character sheaf $\cs{L}=(\gcs{L},\mu,F)$
  on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
    $\EE$-vector spaces $\gcs{L}_x \ceq x^*\gcs{L}$, as $x$ runs over
    $G(\bFq)$;

@@

  \labitem{(cs.2)}{cs.2} an indexed set of isomorphisms $F_{x} \ceq x^*F: \gcs{L}_{\Frob(x)} \to \gcs{L}_x$
    \Clifton{\textdbend There's hidden subtlety here too: maybe I
      should explain which Frobenius this is and how it acts. }
    such that
    \[
    \begin{tikzcd}[row sep=40]
      \gcs{L}_{\Frob(x)+\Frob(y)} \arrow[swap]{d}{F_{x+y}} \arrow{rr}{\mu_{\Frob(x),\Frob(y)}}
      && \gcs{L}_{\Frob(x)}\otimes \gcs{L}_{\Frob(y)} \dar{F_x \tight{0}{\otimes}{0} F_y} \\
      \gcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
      && \gcs{L}_x \otimes\gcs{L}_y
    \end{tikzcd}
    \]
    commutes, for all $x,y\in G(\bFq)$.
  \end{enumerate}
  Under this equivalence, a map $\varphi : \cs{L} \to \cs{L'}$ of
  character sheaves on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.3)}{cs.3} an indexed set $\varphi_x \ceq x^*\varphi: \gcs{L}_x \to \gcs{L'}_x$
    of linear transformations (of one-dimensional $\EE$-vector spaces),
    as $x$ runs over $G(\bFq)$, such that
    \[
    \begin{tikzcd}[column sep=40]
      \arrow[swap]{d}{F_x} \gcs{L}_{\Frob(x)} \rar{\varphi_{\Frob(x)}} & \gcs{L'}_{\Frob(x)} \dar{F_x'}
      &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}& \arrow[swap]{d}{\mu} \gcs{L}_{x+y} \rar{\varphi_{x+y}} & \gcs{L'}_{x+y} \dar{\mu'_{x+y}} \\
      \gcs{L}_x \rar{\varphi_x} & \gcs{L'}_x
      & {} & \gcs{L}_x\otimes\gcs{L}_y \rar{\varphi_x\otimes \varphi_y} & \gcs{L'}_x \otimes \gcs{L'}_y
    \end{tikzcd}
    \]
    both commute.
  \end{enumerate}
  It is clear that the category defined by conditions \ref{cs.0},
  \ref{cs.1}, \ref{cs.2} and \ref{cs.3}, denoted by
  \cdef{$\CS(G(\bFq))$}, is equivalent to $\CS(G)$. We will prove the
  proposition by exhibiting an isomorphism between
  $\CSiso{G(\bFq)}$ and $\Hh^1(W_{\Fq}, G(\bFq)^*)$.

  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(W_{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$. We
  will show that the group isomorphism does not depend on these
  choices.

@@

  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(W_\Fq,G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $F_x : \gcs{L}_{\Frob(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $F_x(s(\Frob(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $t_\cs{L}(\Frob)(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad F_{x}(s(\Frob(x)) = t_\cs{L}(\Frob)(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob)(x)$ is non-zero, while condition~\ref{cs.1} forces
  $t_\cs{L}(\Frob)(x+y) = t_\cs{L}(\Frob)(x) \ t_\cs{L}(\Frob)(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

  Now $t_\cs{L} : W_\Fq \to G(\bFq)^*$ is defined, recursively, by
  $t_\cs{L}(\Frob^n) \ceq t_\cs{L}(\Frob)\cdot \,^\Frob t_\cs{L}(\Frob^{n-1})$,
  where $\Frob$ acts on $G(\bFq)^*$ through pre-composition and $\cdot$ refers
  to pointwise multiplication of functions. This gives
  \[
  \obj \CS(G(\bFq)) \longrightarrow Z^1(W_\Fq,G(\bFq)^*),
  \qquad \text{defined by} \qquad \cs{L} \mapsto t_\cs{L}.
  \]

  Although we used a global section $s_\cs{L}$ in the definition of
  $t_\cs{L}$, the class of $t_\cs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  the line bundle $(\gcs{L},\mu)$, in the sense above and let $t_1$
  and $t_2$ be the cocycles defined by $s_1$ and $s_2$, respectively,
  as above. Since, for each $x\in G(\bFq)$, $\gcs{L}_x$ is a
  one-dimensional $\EE$-vector space, there is a unique scalar
  $a(x) \in \EE$ such that $s_2(x) = a(x) s_1(x)$; moreover, $a(x)$ is
  non-zero, since $s_2(x)$ is non-zero. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$. Now consider the coboundary $t_0 : W_\Fq \to G(\bFq)^*$
  defined by $t_0(\Frob) = \,^\Frob a \cdot a^{-1}$. Then $t_2 = t_0 t_1$,
  so $t_2$ and $t_1$ are cohomologous. In this way we see that
  \begin{equation}\label{cohomologous}
    \obj \CS(G(\bFq)) \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*),
    \qquad \text{defined by} \qquad \cs{L} \mapsto [t_\cs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

  We now show that the function $\obj \CS(G(\bFq)) \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*)$
  factors through $\obj \CS(G(\bFq)) \to \CSiso{G(\bFq)}$.
  To see this, let $\varphi : \cs{L} \to \cs{L'}$ be an isomorphism and
  let $t$ and $t'$ be any cocycles representing the classes in the
  image of $\cs{L}$ and $\cs{L'}$ under \eqref{cohomologous}; by construction,
  $t$ and $t'$ are defined by sections $s$ and $s'$. Now, $\varphi_x(s(x))$ is a
  non-zero scalar multiple of $s'(x)$; define $\phi: G(\bFq) \to \EEx$
  by $\varphi_x(s(x)) = \phi(x) s'(x)$. By the second part of condition~\ref{cs.3},
  $\phi(x+y) = \phi(x)\phi(y)$, so $\phi \in G(\bFq)^*$. By the first part of
  condition~\ref{cs.3}, $t'(\Frob) = t(\Frob) \cdot (\,^{\Frob^{-1}}\phi \cdot \phi^{-1})$,
  so $t$ and $t'$ are cohomologous. This concludes the definition of
  \begin{equation}\label{pich1}
    \CSiso{G(\bFq} \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*),
    \qquad \text{defined by} \qquad [\cs{L}] \mapsto [t_\cs{L}].
  \end{equation}
  Now it follows from \eqref{t} that the image of the isomorphism
  class of $\cs{L}$ is indeed represented by a cocycle $t_\cs{L}$ such
  that $t_\cs{L}(\Frob) = t^\Frob_\cs{L}$ on $G(\Fq)$, which is to
  say, when evaluated at $\Frob$, the cocycle $t_\cs{L}$ is an
  extension of the trace of Frobenius of $\cs{L}$.

@@

%\input{etale-2}
%Define $\rho \colon \Hh^1(W_{\Fq}, \Hom(G(\bFq), \EEx))
% \rightarrow \Hom(G(\Fq), \EEx)$ by $\rho(\phi)(g) =
% \phi(\Frob)(g)$.

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\cs{L}_{k'} = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob_G^*(F) \circ \cdots \circ (\Frob_G^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob_{G_{k'}} = \Frob_G^n$.

\begin{proposition}
The diagram
\[
\begin{tikzcd}
\CSiso{G} \rar \dar & \CSiso{G_{k'}} \dar \\
\Hom(G(k), \EEx) \rar & \Hom(G(k'), \EEx)
\end{tikzcd}
\]
commutes, where vertical arrows are as in Theorem \ref{thm:snake}, the top arrow is as above,
and the bottom is induced by the norm map $G(k') \to G(k)$.
\end{proposition}

>>>>

\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\W}{W}
\newcommand{\Frob}{{\operatorname{Fr}}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\Irrep}{Irrep}
\DeclareMathOperator{\Pic}{Pic}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Hh}{H}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Nm}{Nm}

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}

@@

\begin{remark} \label{rem:CS-action}
If $W$ is a group equipped with a homomorphism $\alpha : W \to \Gal(\bFq/\Fq)$
then an \emph{action} of $W$ on $\gcs{L}$ is a collection of isomorphisms
\[
\gamma_w : \alpha(w)^* \gcs{L} \to \gcs{L}
\]
so that $\gamma_{uw} = \gamma_w \circ \alpha(w)^*\gamma_u$. \David{make sure I have the order right and add a reference}
Specifying an $F$ as in \ref{CS.2} is the same as giving an action of the Weil group $\Weil{k}$
on $\gcs{L}$ compatible with $\mu$.
\end{remark}

@@

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\W(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\W(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil{\Fq}\to \W(G,\bg)$ of the exact sequence
  \[
  \begin{tikzcd}
  1 \rar & \pi_1(\bG, \bg) \rar &  \W(G,\bg) \rar
  & \ar[swap,bend right, dotted]{l}{g} \Weil{\Fq} \rar & 1
  \end{tikzcd}
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil{\Fq} \to \EEx \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil{\Fq}$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = t^\Frob_{(\gcs{L},F)}(g)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil{\Fq}$ is generated by
  $\Frob$ and $t_{(\gcs{L},F)}(g) : \Weil{\Fq} \to \EEx$ is
  a character, it follows that the image of $t_{(\gcs{L},F)}(g)$ is a
  finite subset of $\mu(\EE)$. Thus, $t_{(\gcs{L},F)}(g)$ extends to
  an $\ell$-adic character of $\Gal(\bFq/\Fq)$, which will be denoted
  by the same symbol.
  \[
  \begin{tikzcd}
  1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \W(G,\bg) \rar \dar & \Weil{\Fq} \rar \dar & 1\\
  1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1
  \end{tikzcd}
  \]

@@

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil{\Fq}, \Hom_\text{grp}(G(\bFq),\EEx)),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [t_\cs{L}],
  \]
  where $t_\cs{L}: \Weil{\Fq}\to \Hom_\text{grp}(G(\bFq),\EEx)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad t_\cs{L}(\Frob)(g) = t^\Frob_\cs{L}(g).
  \]
\end{proposition}

\begin{proof}
  To simplify notation, we set
  $G(\bFq)^* \ceq \Hom_\text{grp}(G(\bFq),\EEx)$.  Recall the
  equivalence between etale group schemes over $\Fq$ and the category
  of groups equipped with a continuous action of $\Gal(\bFq/\Fq)$ (for
  the discrete topology on the group), given by $G \mapsto G(\bFq)$.
  Under this equivalence any character sheaf $\cs{L}=(\gcs{L},\mu,F)$
  on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
    $\EE$-vector spaces $\gcs{L}_x \ceq x^*\gcs{L}$, as $x$ runs over
    $G(\bFq)$;

@@

  \labitem{(cs.2)}{cs.2} an indexed set of isomorphisms $F_{x} \ceq x^*F: \gcs{L}_{\Frob(x)} \to \gcs{L}_x$
    \Clifton{\textdbend There's hidden subtlety here too: maybe I
      should explain which Frobenius this is and how it acts. }
    such that
    \[
    \begin{tikzcd}[row sep=40]
      \gcs{L}_{\Frob(x)+\Frob(y)} \arrow[swap]{d}{F_{x+y}} \arrow{rr}{\mu_{\Frob(x),\Frob(y)}}
      && \gcs{L}_{\Frob(x)}\otimes \gcs{L}_{\Frob(y)} \dar{F_x \tight{0}{\otimes}{0} F_y} \\
      \gcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
      && \gcs{L}_x \otimes\gcs{L}_y
    \end{tikzcd}
    \]
    commutes, for all $x,y\in G(\bFq)$.
  \end{enumerate}
  Under this equivalence, a map $\varphi : \cs{L} \to \cs{L'}$ of
  character sheaves on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.3)}{cs.3} an indexed set $\varphi_x \ceq x^*\varphi: \gcs{L}_x \to \gcs{L'}_x$
    of linear transformations (of one-dimensional $\EE$-vector spaces),
    as $x$ runs over $G(\bFq)$, such that
    \[
    \begin{tikzcd}[column sep=40]
      \arrow[swap]{d}{F_x} \gcs{L}_{\Frob(x)} \rar{\varphi_{\Frob(x)}} & \gcs{L'}_{\Frob(x)} \dar{F_x'}
      &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}& \arrow[swap]{d}{\mu} \gcs{L}_{x+y} \rar{\varphi_{x+y}} & \gcs{L'}_{x+y} \dar{\mu'_{x+y}} \\
      \gcs{L}_x \rar{\varphi_x} & \gcs{L'}_x
      & {} & \gcs{L}_x\otimes\gcs{L}_y \rar{\varphi_x\otimes \varphi_y} & \gcs{L'}_x \otimes \gcs{L'}_y
    \end{tikzcd}
    \]
    both commute.
  \end{enumerate}
  It is clear that the category defined by conditions \ref{cs.0},
  \ref{cs.1}, \ref{cs.2} and \ref{cs.3}, denoted by
  \cdef{$\CS(G(\bFq))$}, is equivalent to $\CS(G)$. We will prove the
  proposition by exhibiting an isomorphism between
  $\CSiso{G(\bFq)}$ and $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$.

  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$. We
  will show that the group isomorphism does not depend on these
  choices.

@@

  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $F_x : \gcs{L}_{\Frob(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $F_x(s(\Frob(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $t_\cs{L}(\Frob)(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad F_{x}(s(\Frob(x)) = t_\cs{L}(\Frob)(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob)(x)$ is non-zero, while condition~\ref{cs.1} forces
  $t_\cs{L}(\Frob)(x+y) = t_\cs{L}(\Frob)(x) \ t_\cs{L}(\Frob)(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

  Now $t_\cs{L} : \Weil{\Fq} \to G(\bFq)^*$ is defined, recursively, by
  $t_\cs{L}(\Frob^n) \ceq t_\cs{L}(\Frob)\cdot \,^\Frob t_\cs{L}(\Frob^{n-1})$,
  where $\Frob$ acts on $G(\bFq)^*$ through pre-composition and $\cdot$ refers
  to pointwise multiplication of functions. This gives
  \[
  \obj \CS(G(\bFq)) \longrightarrow Z^1(\Weil{\Fq},G(\bFq)^*),
  \qquad \text{defined by} \qquad \cs{L} \mapsto t_\cs{L}.
  \]

  Although we used a global section $s_\cs{L}$ in the definition of
  $t_\cs{L}$, the class of $t_\cs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  the line bundle $(\gcs{L},\mu)$, in the sense above and let $t_1$
  and $t_2$ be the cocycles defined by $s_1$ and $s_2$, respectively,
  as above. Since, for each $x\in G(\bFq)$, $\gcs{L}_x$ is a
  one-dimensional $\EE$-vector space, there is a unique scalar
  $a(x) \in \EE$ such that $s_2(x) = a(x) s_1(x)$; moreover, $a(x)$ is
  non-zero, since $s_2(x)$ is non-zero. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$. Now consider the coboundary $t_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $t_0(\Frob) = \,^\Frob a \cdot a^{-1}$. Then $t_2 = t_0 t_1$,
  so $t_2$ and $t_1$ are cohomologous. In this way we see that
  \begin{equation}\label{cohomologous}
    \obj \CS(G(\bFq)) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \cs{L} \mapsto [t_\cs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

  We now show that the function $\obj \CS(G(\bFq)) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*)$
  factors through $\obj \CS(G(\bFq)) \to \CSiso{G(\bFq)}$.
  To see this, let $\varphi : \cs{L} \to \cs{L'}$ be an isomorphism and
  let $t$ and $t'$ be any cocycles representing the classes in the
  image of $\cs{L}$ and $\cs{L'}$ under \eqref{cohomologous}; by construction,
  $t$ and $t'$ are defined by sections $s$ and $s'$. Now, $\varphi_x(s(x))$ is a
  non-zero scalar multiple of $s'(x)$; define $\phi: G(\bFq) \to \EEx$
  by $\varphi_x(s(x)) = \phi(x) s'(x)$. By the second part of condition~\ref{cs.3},
  $\phi(x+y) = \phi(x)\phi(y)$, so $\phi \in G(\bFq)^*$. By the first part of
  condition~\ref{cs.3}, $t'(\Frob) = t(\Frob) \cdot (\,^{\Frob^{-1}}\phi \cdot \phi^{-1})$,
  so $t$ and $t'$ are cohomologous. This concludes the definition of
  \begin{equation}\label{pich1}
    \CSiso{G(\bFq} \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad [\cs{L}] \mapsto [t_\cs{L}].
  \end{equation}
  Now it follows from \eqref{t} that the image of the isomorphism
  class of $\cs{L}$ is indeed represented by a cocycle $t_\cs{L}$ such
  that $t_\cs{L}(\Frob) = t^\Frob_\cs{L}$ on $G(\Fq)$, which is to
  say, when evaluated at $\Frob$, the cocycle $t_\cs{L}$ is an
  extension of the trace of Frobenius of $\cs{L}$.

@@

%\input{etale-2}
%Define $\rho \colon \Hh^1(\Weil{\Fq}, \Hom(G(\bFq), \EEx))
% \rightarrow \Hom(G(\Fq), \EEx)$ by $\rho(\phi)(g) =
% \phi(\Frob)(g)$.

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob_G^*(F) \circ \cdots \circ (\Frob_G^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob_{G_{k'}} = \Frob_G^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\CSiso{G} \rar{\nu} \dar{t_G} & \CSiso{G_{k'}} \dar{t_{G_{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}
